---
title: "DArT workflow"
author: "Christian Bowman"
output: pdf_document
---

## Dependencies & Information
R 3.6.3 \newline
Rstudio 1.3.1056 \newline
Ubuntu 20.04.2 LTS \newline
dartR_1.8.3 \newline
StAMPP_1.6.1 \newline

```{r setup, echo=FALSE}
knitr::opts_knit$set(root.dir = "/home/cb/workspace")
pacman::p_load(dartR, StAMPP)
source("gl.filter.PIC.r")
source("gl2structure_localfix.r")

```

# 2017 analysis

## Working variables - 2017 data
```{r vars}
setwd("./bermuda/2017-12-06")

nfile <- "Report_DBe17-3222_SNP_2.csv"
sfile <- "Report_DBe17-SilicoDArT_1.csv"
mdata <- "Meta_DBe17-3222.csv"

############################################################################
## Setting light switch for easy swap between SNP(0) & Silico(1) analysis ##
flick <- 0
############################################################################

indivs <- c("B1","B99","B106","B114","B17","B27","B91","B2","B100","B116","B9","B63","B108","B117","B19","B30","B38","B65","B4","B101","B109","B119","B11","B20","B31","B39","B56","B66","B94","B5","B102","B110","B120","B12","B21","B40","B50","B95","B6","B121","B13","B42","B96","B7","B104","B112","B14","B43","B51","B60","B97","B8","B16","B44","B52","B61","B98","B140","B147","B160","B133","B141","B148","B162","B125","B134","B142","B149","B135","B151","B127","B136","B144","B152","B128","B137","B145","B156","B129","B138","B146","B158","B131","B139","B159")

if (flick == 0) {
  cat("SNP data detected")
  phold <- nfile
  gl <- gl.read.dart(filename = nfile, ind.metafile = mdata)
  gl.i <- gl <- gl.keep.ind(gl, ind.list=indivs)
  gl.0 <- gl.recalc.metrics(gl.i)
  
} else if (flick == 1) {
  cat("Silico data detected")
  phold <- sfile
  gl <- gl.read.silicodart(filename = sfile, ind.metafile = mdata)
  gl.i <- gl <- gl.keep.ind(gl, ind.list=indivs)
  gl.0 <- gl.recalc.metrics(gl.i)  

} else {
  cat("Error: data type not specified (0 = SNP, 1 = Silico)")

}

```

## Reporting & Filtering - 2017 data

2017 SNP : Pre-report
```{r pre-report}
gl.report.reproducibility(gl.0)
gl.report.monomorphs(gl.0)
gl.report.callrate(gl.0)
gl.report.secondaries(gl.0)

```

2017 SNP : Filter
\begin{itemize}
  \item (-) reproducibility < 1
  \item (-) monomorphic markers
  \item (-) call rate < 95%
  \item (-) secondaries
\end{itemize}

2017 silico : Filter
\begin{itemize}
  \item (-) reproducibility < 1
  \item (-) monomorphic markers
  \item (-) call rate < 95%
  \item (-) PIC < 0.4
\end{itemize}

```{r filter}
if (flick == 0) {
  ##SNP
  glf.1 <- gl.filter.reproducibility(gl.0, threshold = 1)    #quality per locus; .repavg is deprecated 
  glf.2 <- gl.filter.monomorphs(glf.1)    #remove monomorphs and loci w/ NA
  glf.3 <- gl.filter.callrate(glf.2, method = "loc", threshold = 0.95)    #missing data per locus
  glf.4 <- gl.filter.secondaries(glf.3)    #remove dupes per seq tag  
  
} else if (flick == 1) {
  ##Silico
  glf.1 <- gl.filter.reproducibility(gl.0, threshold = 1)    #quality per locus
  glf.2 <- gl.filter.monomorphs(glf.1)    #remove monomorphs and loci w/ NA
  glf.3 <- gl.filter.callrate(glf.2, method = "loc", threshold = 0.95)    #missing data per locus
  glf.4 <- gl.filter.PIC(glf.3, threshold = 0.4)    #polymorphic information content
  
}

rm(glf.1, glf.2, glf.3)
```

2017 SNP : Post-report
```{r post-report}
if (flick == 0) {
  ##SNP
  gl.report.reproducibility(glf.4)
  gl.report.monomorphs(glf.4)
  gl.report.callrate(glf.4)
  gl.report.secondaries(glf.4)
    
} else if (flick == 1) {
  ##Silico
  gl.report.reproducibility(glf.4)
  gl.report.monomorphs(glf.4)
  gl.report.callrate(glf.4)
  #gl.report.(glf.4)
  
}

```

## Measuring Genetic Distance - 2017 data

```{r Genetic distance}
##F statistic; Bootstrap=1000
pwfst <- stamppFst(glf.4, nboots=1000, percent=95, nclusters=2)
write.csv(pwfst$Fsts,file=paste0("FST_",phold))
write.csv(pwfst$Pvalues,file=paste0("FSTpval_",phold))

##G statistic (Neis genetic distance)
pwGst <- stamppNeisD(glf.4)
round(pwGst,3)
write.csv(pwGst,file=paste0("GST_",phold))

##G statistic for indivs (Nei genetic distance for individuals)
glind <- glf.4
pop(glind) <- indNames(glind)
pwGst.i <- stamppNeisD(glind)
round(pwGst.i,3)
write.csv(pwGst.i,file=paste0("GSTindiv_",phold))

```
Couldn't generate Fst for individuals, I think this method is only for populations.

Files were exported as csv files.

## PCoA analysis - 2017 data

Building PCoA with white background
```{r PCoA}
#Need to clean up the PCoA graph (visuals)
#glind<-glf.4
pc <- gl.pcoa(glind, nfactors=12)    #explanation and results for PCoA
y <- gl.pcoa.plot(pc, glind, labels="ind", xaxis=1, yaxis=2)
y + xlim(-25, 25)+ylim(-25, 25)+coord_fixed(ratio = 1)+theme_bw()+geom_point(aes(shape=pop))
#geom_point works only for 6 shapes.

#Scree plot with eigenvalues to determine contribution of axes in explaining variation
gl.pcoa.scree(pc)


```
Output:
Performing a PCoA, individuals as entities, SNP loci as attributes
Ordination yielded 15 informative dimensions from 84 original dimensions
  PCoA Axis 1 explains 22.7 % of the total variance
  PCoA Axis 1 and 2 combined explain 35.6 % of the total variance
  PCoA Axis 1-3 combined explain 41.7 % of the total variance

## STRUCTURE Export
```{r STRUCTURE}
# does STRUCTURE only handle SNP markers or can it handle Silico?
gl2structure_localfix(glf.4)

```

## PHYLIP Export
```{r PHYLIP}
### Calculate Nei's distances between individuals/pops
#how does this compare with the final lines of ##Measuring Genetic distance ?
aa.D.ind <- stamppNeisD(glind, pop = FALSE) # Nei's 1972 distance between indivs
stamppPhylip(aa.D.ind, file="bermudasilico.indiv_Neis_distance.txt") # export matrix - for SplitsTree
aa.D.pop <- stamppNeisD(aa.genlight, pop = TRUE) # Nei's 1972 distance between pops
stamppPhylip(aa.D.pop, file="aa.pops_Neis_distance.phy.dst") # export matrix - for SplitsTree 

nLoc(glind)

```

## Outputs - 2017 data
###\underline{SNP}
Raw: 
\begin{itemize}
  \item 85 individuals
  \item 192504 (178730 in paper) SNPS
  \item (40.99%) missing data
\end{itemize}

Filtered - SNP:
\begin{itemize}
  \item 85 individuals
  \item 7023 (paper match) SNPs
  \item (2.04%) missing data
\end{itemize}

###\underline{silico}
Raw:
\begin{itemize}
  \item 85 individuals
  \item 376269 (285570) SNPS
  \item (1.84%) missing data
\end{itemize}

Filtered - Silico:
\begin{itemize}
  \item 85 individuals
  \item 27512 (30505) SNPs
  \item (3.17%) missing data
\end{itemize}

# 2018 analysis

# Notes - 2018 data

\underline{Metadata}
\begin{itemize}
  \item created "Meta_DBe18-3712.csv"
  \item "Genotype" -> "id"
  \item "Species" -> "pop"
  \item "Cynodon <species>" -> "C.<species>"
  \item "Comments" -> "epop" (expected population)
  \item "\textbackslash xD7" -> "x" (\textbackslash xD7 is meant to be a multiplication sign)
  \item "Comments: none" -> "N/A"; not sure how missing values are interpreted in the metadata
  \item [FIXED] ID:"FA8-1 " -> ID:"FA8-1"; fixed in raw and meta
\end{itemize}

\underline{Silico}
\begin{itemize}
  \item must use read.silicodart
  \item removing secondaries is not yet implemented; must use PIC filter <0.4
  \item created "gl.filter.PIC" using "gl.filter.reproducibility" as the base; filter based off of PIC t=0.4
  \item 
  

